---
title: "nutrients"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```

Goal: combine nutrient data (serc and ws) per reef and subregion.

## read data

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(DT)

tbl_serc <- read_excel("data/Bird_Key_Reef_serc.xls")
tbl_ws   <- read_excel("data/Bird_Key_Reef_ws.xls")
```

## show data

Notice how `tbl_serc.DATE_` and `tbl_ws.Date_`, displayed as tibbles, are already type `dttm`, ie datetime. And that these dates are not already aggregated by year.

To look at these tables in RStudio, use in the Console `View(tbl_serc)`. The `View()` function is only available for interactive use.


```{r}
tbl_serc
```

```{r}
tbl_ws
```

What are fields 11 through 16 after `Phaeophyti` in `tbl_ws`?

## summarize by year

Before we join by year we have to aggregate by year, and decide what fields will we additionally join by?

Per the original goal "to combine nutrient data (serc and ws) per reef and subregion", here are potentially valid fields:

* tbl_serc: BASIN, SEGMENT, ZONE_, STATION, SITE, [LATDEC, LONDEC]
* tbl_ws: Station, [Longitude, Latitude]

It looks like tbl_serc.STATION (dbl) and tbl_ws.Station (chr) are the only common fields, but they are different data types: double (dbl) vs character (chr). So how do you relate reef and subregion in tbl_serc and especially tbl_ws based on given fields?

How do you want to summarize multiple values for a given year? Let's simply take the average to start.

```{r}
tbl_yr_chl <- tbl_serc %>% 
  select(date = DATE_, chla = CHLA_S) %>% 
  bind_rows(
    tbl_ws %>% 
      select(date = Date_, chla = Chl_a)) %>% 
  mutate(
    yr = year(date)) %>% 
  group_by(yr) %>% 
  summarize(chla = mean(chla))

datatable(tbl_yr_chl)
```


